EEG characterization and classification based on the histograms of visual gradients of plots

### 2016-12-18 article.tex
General layout of the paper

Cybathlon

x) El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
x) La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
x) Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
x) La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
x) La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
x) La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
x) Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
x) ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).

@@

Their versatility, ease of use, temporal resolution, ease and development of fabrication, and the explosion as a consumer devices, is pushing EEG devices to become the de facto non invasive method to access and harness brain information.

@@

One very remarkable aspect of this communication channel is the ability to transmit some general cognitive state, like alertness, drowsiness, boredom, and so on, which can very helpful particularly in rehabilitation procedures. \citep{Guger2009a}.

One of the most shocking reasons is that we can use the capacity of ML techniques to automatically explore the signal, by using these new features.

Additionally, they have a high variability between different subjects and even between different moments for the same subject.  This inherent complexity comes from  is a real challenge when it is required to extract information from raw EEG signals.

Signal is complex and it is difficult to find those values which are not variable.

Due to this inner complexity, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, classify it, and try to determine some meaning out of it.

In this work EEG signals from ALS patients are offline processed with this radically different and horizontal approach.  This is the continuation of the work presented here and here \citep{Ramele2016}, which was previously applied to rythmic patterns.

@@

Successful applications of BCI has been applied to rehabilitation \citep{Jure2016}.

@@

BCI accuracy
was calculated as the percentage of attempted selections
that matched the intended selection.

@@

The protocol is explained [HERE] but can be summarized as follows:  8 subjects affected by different levels of ALS disease, were used to generate all the data. The P300 task was meant to spell 7 words of 5 letter each.  The first 3 runs were used for training and the remaining 4 for testing with visual feedback.  A trial, for this experiment, was defined as each attempt to select a letter from the speller, and it was composed of 10 repetions of flashes of 6 rows and 6 columns of the matrix, yielding 120 repetitions.  Flashing of row/columns was performed for 0.125 s, following by a resting period of the same size.

This dataset was recorded from a healthy subject during a no-feedback session. She/he sat in a normal chair with relaxed arms resting on the table and fingers in the standard typing position at the computer keyboard. The task was to press with the index and little fingers the corresponding keys in a self-chosen order and timing 'self-paced key typing'. The experiment consisted of 3 sessions of 6 minutes each. All sessions were conducted on the same day with some minutes break in-between. Typing was done at an average speed of 1 key per second.

@@

Due to the inner complexity of EEG signals, several approaches and methods has been applied to decode them.  Of particular interests are morphological signal analysis 
Paper del chabon del ITBA que detalla bien los diferentes metodos,

paper de los mexicanos \citep{Alvarado-Gonzalez2016}.

@@

Results are shown in the following chart.

\begin{figure}[H]
\centering
\includegraphics[width=18cm]{subjectaverageds4.eps}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

Contrary to the literature (\citep{Huggins2016,Jure2016}) the best performing channels were not Cz and Pz and instead Pz.

GRAND AVERAGES

it is quite likely that the data will be unbalanced (that is, an unequal number of trials in each condition)

\begin{table}[H]
\caption{Results obtained for Accuracy levels by k-fold cross validation.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{cccc}
\toprule
\textbf{Cz}	& \textbf{Cz}	& \textbf{Best Channel}	& \textbf{Performance}\\
\midrule
0.845000     & 0.794286 & 2 & 0.794286 $\pm$ 0.011952 \\
0.863000     & 0.811429 & 7 & 0.934286 $\pm$ 0.011200 \\
0.872000     & 0.780000 & 2 & 0.780000 $\pm$ 0.031843 \\
0.859000     & 0.675714 & 8 & 0.904286 $\pm$ 0.013232 \\
0.862000     & 0.801429 & 7 & 0.932857 $\pm$ 0.008400 \\
0.886000     & 0.964286 & 2 & 0.964286 $\pm$ 0.010435 \\
0.886000     & 0.784286 & 7 & 0.930000 $\pm$ 0.009744 \\
0.923000     & 0.908571 & 7 & 0.997143 $\pm$ 0.003733 \\
\bottomrule
\end{tabular}
\end{table}

@@

Results of one against the other

@@

Check if you can use this to detect any kind of transient signbal

>>>>

\begin{enumerate}[leftmargin=*,labelsep=3mm]
\item El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
\item La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
\item Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
\item La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
\item La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
\item La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
\item Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
\item ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
\end{enumerate}

@@

Their versatility, ease of use, temporal resolution, ease and development of fabrication, and the explosion as a consumer devices, is pushing EEG devices to become the de-facto non invasive method to access and harness brain information.

@@

One very remarkable aspect of this communication channel is the ability to volitionally transmit information from the CNS, Central Nervous System, to a computer device and from there use that information to drive wheelchair or as input in a speller application \citep{Guger2009a}, or a rehabilitation procedure \citep{Jure2016} in connection to FES.

EEG signals are remarkably complex and has been characterized as a non-stationary stochastic process with even sometimes their baseline activity is near random (CITE TEODIANO BASTOS).

Additionally, they have a high variability between different subjects and even between different moments for the same subject.  This inherent complexity is a real challenge when it is required to extract information from raw EEG signals.

Hence, inevitably, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.

In this work a new method to characterize EEG signals is presented, expanded and detailed.  Its validity is shown by processing offline data from ALS patients.  This is the continuation of the work previously presented \citep{Ramele2016}, which was applied mainly to rythmic patterns.

@@

@@

@@

The protocol is explained \citep{Riccio2013} but can be summarized as follows:  8 subjects affected by different levels of ALS disease, were used to generate all the data. The P300 task was meant to spell 7 words of 5 letter each.  The first 3 runs were used for training and the remaining 4 for testing with visual feedback.  A trial, for this experiment, was defined as each attempt to select a letter from the speller, and it was composed of 10 repetitions of flashes of 6 rows and 6 columns of the matrix, yielding 120 repetitions.  Flashing of row/columns was performed for 0.125 s, following by a resting period of the same size.  After 120 repetions and intertrial pause is performed to start with the following letter.

All sessions were conducted on the same day with some minutes break in-between. Typing was done at an average speed of 1 key per second.

@@

Due to the inner complexity of EEG signals, several approaches and methods have been applied to decode them.  Of particular interests are morphological signal analysis 
CITE (Paper del chabon del ITBA que detalla bien los diferentes metodos) \citep{Alvarado-Gonzalez2016}.

@@

In Fig. \ref{fig:subjectaveraged} the grand average (point-to-point) of for all the subjects is shown.  The P300 characteristic curve can be seen particularly in subjects 6, and 7 and in a lesser extend in the remaining subjects. In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (signal variance is proportional to the root of the number of samples and the procedure would be discriminating signals with different variances).it is quite likely that the data will be unbalanced (that is, an unequal number of trials in each condition)

\begin{figure}[H]
\centering
\includegraphics[width=18cm]{subjectaverageds4.eps}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\label{fig:subjectaveraged}
\end{figure}

Results are shown in the following chart where BCI accuracy
was calculated as the percentage of successful predictions
that matched the proposed selection in a copy-spelling approach (CITE TAYLOR AND  FRANCOS PAPER).

\begin{table}[H]
\caption{Results obtained for Accuracy levels by k-fold cross validation. The values reported for Cz are reproduced here for comparison. Cz values of BCI accuracies can be seen as long as the accuracy levels obtained for the best performing channels, for each subject.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{cccc}
\toprule
\textbf{Cz}	& \textbf{Cz}	& \textbf{Best Channel}	& \textbf{Performance}\\
\midrule
0.845000     & 0.794286 & 2 & 0.794286 $\pm$ 0.011952 \\
0.863000     & 0.811429 & 7 & 0.934286 $\pm$ 0.011200 \\
0.872000     & 0.780000 & 2 & 0.780000 $\pm$ 0.031843 \\
0.859000     & 0.675714 & 8 & 0.904286 $\pm$ 0.013232 \\
0.862000     & 0.801429 & 7 & 0.932857 $\pm$ 0.008400 \\
0.886000     & 0.964286 & 2 & 0.964286 $\pm$ 0.010435 \\
0.886000     & 0.784286 & 7 & 0.930000 $\pm$ 0.009744 \\
0.923000     & 0.908571 & 7 & 0.997143 $\pm$ 0.003733 \\
\bottomrule
\end{tabular}
\end{table}

Contrary to the literature (\citep{Huggins2016,Jure2016}) the best performing channels were not Cz and Pz but instead the occipital channels PO1 and P02 (5 against 3).

@@

@@

Check if you can use this to detect any kind of transient signal.

Compare if it is possible with the descriptors from one subject, discriminate the others.

Channal identification based on the metric distance between the bags
